A Short and Mechanized Consistency Proof for Dependent Type Theory

### 2024-02-15 paper/paper.tex
Tweak the notation for renamign and substitution

The syntax of \lang can be found in Figure~\ref{fig:syntax}. We use
the unscoped de Bruijn representation for both our Coq development and
the informal presentation in our paper. We find de Bruijn
representation advantageous for specification since it leaves very
little ambiguity about the variable freeness side
conditions, making our proof more easily reproducible. Furthermore, as we discuss in
Section~\ref{sec:automation}, de Bruijn representation is much more
amenable to automated reasoning in proof assistants.
We adopt from \citet{autosubst} the notations for simultaneous
renaming, substitution, and other auxiliary definitions related to our
syntax, summarized in Figure~\ref{fig:auxdef}.

\begin{figure}[ht]
  \begin{equation*}
    \begin{split}
      \begin{array}{lll}
        \mathit{Identity (Ren)} \\
        [[id i]] & := & [[i]] \\ \\
        \mathit{Identity (Subst)} \\
        [[idtm i]] & := & [[var i]] \\ \\
        \mathit{Cons (Ren)} \\
        ([[xi .: n]])([[0]]) & := & [[n]] \\
        ([[xi .: n]])([[Suc i]]) & := & [[xi i]] \\ \\
        \mathit{Cons (Subst)} \\
        ([[rho .: a]])([[0]]) & := & [[a]] \\
        ([[rho .: a]])([[Suc i]]) & := & [[rho i]] \\ \\
      \end{array}
    \end{split}
    \qquad \qquad
    \begin{split}
      \begin{array}{lll}
        \mathit{Curried\ Add} \\
        [[up n i]] & := & [[n]] + [[i]] \\ \\
        \mathit{Renaming} \\
        [[a < xi >]] & := & \ldots \\ \\
        \mathit{Up} \\
        [[( up rho ) 0]] & := & [[var 0]] \\
        [[( up rho ) Suc i]] & := & [[ rho i < up 1 > ]] \\ \\
        \mathit{Substitution} \\
        [[var i { rho }  ]] & := &  [[rho i]] \\
        [[(Pi A B) { rho }]] & := & [[Pi A { rho } B { up rho }]] \\
        [[a { rho }  ]] & := & \ldots \\
      \end{array}
    \end{split}
  \end{equation*}
  \caption{Auxiliary Functions over Syntax}
  \label{fig:auxdef}
\end{figure}

>>>>

The syntax of \lang can be found in Figure~\ref{fig:syntax}. We use
the unscoped de Bruijn representation for both our Coq development and
the informal presentation in our paper. We find de Bruijn
representation advantageous for specification since it leaves very
little ambiguity about the variable freeness side
conditions, making our proof more easily reproducible. Furthermore, as we discuss in
Section~\ref{sec:automation}, de Bruijn representation is much more
amenable to automated reasoning in proof assistants.
We adopt from \citet{autosubst} the notations for simultaneous
renaming, substitution, and other auxiliary definitions related to our
syntax, summarized in Figure~\ref{fig:auxdef}. We omit most of the
definitions of renaming and substitution and only show the definition
of a few representative cases of substitution.
\begin{figure}[ht]
  \begin{equation*}
    \begin{split}
      \begin{array}{lll}
        \mathit{IdentityRen} \\
        [[id i]] & := & [[i]] \\ \\
        \mathit{IdentityTm} \\
        [[idtm i]] & := & [[var i]] \\ \\
        \mathit{ConsRen ([[xi .: n]])} \\
        ([[xi .: n]])([[0]]) & := & [[n]] \\
        ([[xi .: n]])([[Suc i]]) & := & [[xi i]] \\ \\
        \mathit{ConsSubst ([[rho .: a]])} \\
        ([[rho .: a]])([[0]]) & := & [[a]] \\
        ([[rho .: a]])([[Suc i]]) & := & [[rho i]] \\ \\
        \mathit{Curried\ Add} \\
        [[up n i]] & := & [[n]] + [[i]]
      \end{array}
    \end{split}
    \qquad \qquad
    \begin{split}
      \begin{array}{lll}
        \mathit{UpRen} \\
        [[( up xi ) 0]] & := & [[0]] \\
        [[( up xi ) Suc i]] & := & [[ Suc ren xi i ]] \\ \\
        \mathit{UpTm} \\
        [[( up rho ) 0]] & := & [[var 0]] \\
        [[( up rho ) Suc i]] & := & [[ rho i < up 1 > ]] \\ \\
        \mathit{Renaming ([[a < xi >]])} \\
        [[(Pi A B) < xi >]] & := & [[Pi A < xi > B < up xi >]] \\
        [[(a b) < xi >]] & := & [[a < xi > (# b < xi > #)]] \\
        \ldots \\ \\
        \mathit{Substitution ([[ a { rho }  ]])} \\
        [[var i { rho }  ]] & := &  [[rho i]] \\
        [[(Pi A B) { rho }]] & := & [[Pi A { rho } B { up rho }]] \\
        [[(a b) { rho }]] & := & [[a { rho } (# b { rho } #)]] \\
        \ldots
      \end{array}
    \end{split}
  \end{equation*}
  \caption{Auxiliary Functions over Syntax}
  \label{fig:auxdef}
\end{figure}
